---
title: "Predictions after COVID-19"
output:
  html_document:
    code_folding: show
    highlight: textmate
    lib_dir: site_libs
    self_contained: yes
    theme: cosmo
    toc: yes
    toc_depth: 5
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '5'
editor_options:
  
  chunk_output_type: console
---

```{r setup, include = F}

knitr::opts_chunk$set(
  warning = F, # show warnings during codebook generation
  message = F, # show messages during codebook generation
  error = F, # do not interrupt codebook generation in case of errors,
                # usually better for debugging
  echo = T,  # show R code
fig.width=12, 
fig.height=8)

options(max.print = 20000, scipen = 100)

```


```{r libraries}

library(haven)
library(dplyr)
library(ggplot2)
library(psych)
library(cowplot)
library(gridExtra)
library(tidyr)
library(plotly)
library(lme4)
library(jtools)
library(forcats)
library(car)

```


```{r import}

setwd("C:\\Users\\igrossma\\Dropbox\\Predictions\\") 

pred.df<-read.csv("predictions.csv")
```


# Data Cleaning

```{r remove.bad.responses}

# Remove missing reposnses for participants who either skipped or did not enter anything for key variables.

# Recode - 99 to Missing
pred.df[pred.df == -99] <- NA

pred.df$filter <- rowSums(pred.df[,c("trust_rating_1", "trust_rating_2", "trust_rating_3", "trust_probability_1", "sat_rating_1", "sat_rating_2", "sat_rating_3", "sat_probability_1", "dep_rating_1", "dep_rating_2", "dep_rating_3", "dep_probability_1", "pol_rating_1", "pol_rating_2", "pol_rating_3", "pol_probability_1", "ind_rating_1", "ind_rating_2", "ind_rating_3", "ind_probability_1", "trad_rating_1", "trad_rating_2", "trad_rating_3", "trad_probability_1", "grat_rating_1", "grat_rating_2", "grat_rating_3", "grat_probability_1", "ex_pre_rating_1", "ex_pre_rating_2", "ex_pre_rating_3", "ex_pre_probability_1", "imp__pre_rating_1", "imp__pre_rating_2", "imp__pre_rating_3", "imp_pre_probability_1", "cli_rating_1", "cli_rating_2", "cli_rating_3", "cli_probability_1", "birth_rating_1", "birth_rating_2", "birth_rating_3", "birth_probability_1")], na.rm = T)

pred.df.clean <- pred.df %>% filter(filter != 0)

# Removed: 668 - 471 = 197 participants who basically just skipped throught the survey.

```

<br>

> 30% of the sample just clicked through and did not even fill out the key parts.

```{r remove.duplicate.responses}

# Create row number variable

pred.df.clean$Row.num <- seq(1, nrow(pred.df.clean), 1)

duplicates.1 <- duplicated(pred.df.clean$IPAddress, fromLast = T)

duplicates.2 <- duplicated(pred.df.clean$IPAddress, fromLast = F)

# pulled out the index of duplicated values

duplicate.rows <- which(duplicates.1 == T | duplicates.2 == T)

# Viewed all duplicated values and made a list below of which ros to drop
#View(pred.df.clean[duplicate.rows, c("IPAddress", "Row.num", "Progress","StartDate", "EndDate")])

# 309 - first incomplete attempt - 20%
# 293 - first incomplete attempt - 17%
# 301 - first incomplete attempt - 30%
# 345 - first incomplete attempt - 55%
# 283 - first incomplete attempt - 27%
# 445 - first incomplete attempt - 30%
# 388 - second complete attempt going with the first.

rows.to.remove <- c(309, 293, 301, 345, 283, 445, 388)

pred.df.clean.x <- filter(pred.df.clean, !(Row.num %in% rows.to.remove))

# removed 7 duplicates.

```

```{r additional.bad.responses}

# Age: 109, 234 and 3254236 - their responses look weird - I don't trust them

age.to.remove <- as.numeric(unlist(pred.df.clean.x[which(pred.df.clean.x$age > 108), "Row.num"]))

pred.df.clean.x.x <- filter(pred.df.clean.x, !(Row.num %in% c(age.to.remove, 412, 190)))

# removed 5 additional participants. - 412: one person didn't enter anything and then entered f*** u for country.
# 190 - test response

```

## Summary

> Started with **`r nrow(pred.df)`** responses. First removed **`r nrow(pred.df) - nrow(pred.df.clean)`** participants because they just clicked through the survey without responding. Then removed another **`r nrow(pred.df.clean) - nrow(pred.df.clean.x)`** duplicate responses. Finally removed **`r nrow(pred.df.clean.x) - nrow(pred.df.clean.x.x)`**. Left with **`r nrow(pred.df.clean.x.x)`** participants.

# Descriptives

```{r all.graphs}

# Age

age.plot <- ggplot(pred.df.clean.x.x, aes(x = age)) + 
  geom_histogram(color = "white") + 
  ggpubr::theme_pubclean(base_size = 16) + 
  theme_minimal(base_size = 16) + 
  theme(axis.line = element_line(color = "gray")) + 
  labs(x = "Age", y = "Frequency") + 
  geom_vline(xintercept = mean(pred.df.clean.x.x$age, na.rm = T), colour = "red", linetype = 2, size = 1)

# Level

level.labels <- c("tenured \nfaculty", "nontenured \nfaculty", "adjunct \nprofessor", "postdoc", "graduate \nstudent", "research \nscientist", "other")

level.plot <- ggplot(pred.df.clean.x.x, aes(x = level)) + 
  geom_histogram(color = "white",  binwidth = 1, boundary = 0.5) + 
  ggpubr::theme_pubclean(base_size = 16) + 
  theme_minimal(base_size = 16) + 
  theme(axis.line = element_line(color = "gray")) + 
  labs(x = "Level", y = "Frequency") + 
  geom_vline(xintercept = mean(pred.df.clean.x.x$level, na.rm = T), colour = "red", linetype = 2, size = 1) + 
  scale_x_continuous(breaks = seq(1, 7, 1), labels = level.labels)

# Unisize

unisize.labels <- c("< 10", "11-100", "101-\n1,000", "1,001-\n10,000", "10,001-\n50,000", "50,000+")

unisize.plot <- ggplot(pred.df.clean.x.x, aes(x = unisize)) + 
  geom_histogram(color = "white",  binwidth = 1, boundary = 0.5) + 
  ggpubr::theme_pubclean(base_size = 16) + 
  theme_minimal(base_size = 16) + 
  theme(axis.line = element_line(color = "gray")) + 
  labs(x = "Unisize", y = "Frequency") + 
  geom_vline(xintercept = mean(pred.df.clean.x.x$unisize, na.rm = T), colour = "red", linetype = 2, size = 1) + 
  scale_x_continuous(breaks = seq(2, 7, 1), labels = unisize.labels)

# Field of research

research.labels <- c("Psychology", "Neuroscience", "Medicine", "Sociology", "Political \nscience", "Economics", "Epidemiology", "Biology", "Computer \nscience", "Other")

research.plot <- ggplot(pred.df.clean.x.x, aes(x = research)) + 
  geom_histogram(color = "white",  binwidth = 1, boundary = 0.5) + 
  ggpubr::theme_pubclean(base_size = 16) + 
  theme_minimal(base_size = 16) + 
  theme(axis.line = element_line(color = "gray")) + 
  labs(x = "Field of Research", y = "Frequency")  + 
  scale_x_continuous(breaks = seq(1, 10, 1), labels = research.labels)

```

## Table

```{r all.descriptives}

knitr::kable(describe(pred.df.clean.x.x[,c("age", "unisize", "level")]), digits = 2)

```

## Graph

```{r all.graphs.combine, fig.width = 12, fig.height=14}

marrangeGrob(
    grobs = list(age.plot, unisize.plot, level.plot, research.plot),
    widths = c(1, 1),
    heights = c(2, 2, 2),
    layout_matrix = rbind(c(1, 2),
                          c(3, 3),
                          c(4, 4)
                          ))
```

## Frequencies

```{r all.frequencies}

table(pred.df.clean.x.x$gender)

table(pred.df.clean.x.x$residence)

```

```{r fix.residence}

pred.df.clean.x.x$residence[pred.df.clean.x.x$residence %in% c("CAN", "canada")] <- "Canada"
pred.df.clean.x.x$residence[pred.df.clean.x.x$residence %in% c("U.S.A", "U.S.A.", "united states", "United States of America", "us", "US", "usa", "Usa", "USA", "USA/Russia", "El Paso", "utah", "United States", "U.S.", "Los Angeles")] <- "Unites States"
pred.df.clean.x.x$residence[pred.df.clean.x.x$residence %in% c("Uk", "United Kingdom", "UK")] <- "United Kingdom"
pred.df.clean.x.x$residence[pred.df.clean.x.x$residence %in% c("germany")] <- "Germany"
pred.df.clean.x.x$residence[pred.df.clean.x.x$residence %in% c("india")] <- "India"
pred.df.clean.x.x$residence[pred.df.clean.x.x$residence %in% c("Hk")] <- "Hong Kong"
pred.df.clean.x.x$residence[pred.df.clean.x.x$residence %in% c("netherlands", "NL", "The Netherlands")] <- "Netherlands"
pred.df.clean.x.x$residence[pred.df.clean.x.x$residence %in% c("sweden")] <- "Sweden"
pred.df.clean.x.x$residence[pred.df.clean.x.x$residence %in% c("switzerland")] <- "Switzerland"
pred.df.clean.x.x$residence[pred.df.clean.x.x$residence %in% c("Turkru")] <- "Turkey"

pred.df.clean.x.x$residence[pred.df.clean.x.x$residence %in% c("Public Affairs", "Riverside")] <- NA

knitr::kable(table(pred.df.clean.x.x$residence))

```


## Confidence

```{r confidence}

knitr::kable(describe(pred.df.clean.x.x[,c(grep("probability_1", names(pred.df.clean.x.x)))]), digits = 2)

```

## Summary

> * Average age is about 40. <br>
> * Gender is close to 50 - 50 split. <br>
> * Most of the participants are either faculty members or grad students in Psychology. Not enough variance to do comparisons with other fields. However, could do graduate students and postdocs vs. tenured and non-tenured faculty. <br>
> * Most participants come from small to large universities in Canada and North America. If we want to use this variable it would need to be fixed as people spelled countried differently.
> * Everyone is from college/university.




# Visualize

```{r convert.to.lng}

# Rename some vars for consistency

colnames(pred.df.clean.x.x)[grep("pre_rating", names(pred.df.clean.x.x))] <- c("ex_rating_1", "ex_rating_2", "ex_rating_3", "imp_rating_1", "imp_rating_2", "imp_rating_3")

#add zero points


pred.df.clean.x.x$cli_rating_0<-0
pred.df.clean.x.x$birth_rating_0<-0
pred.df.clean.x.x$grat_rating_0<-0
pred.df.clean.x.x$dep_rating_0<-0
pred.df.clean.x.x$sat_rating_0<-0
pred.df.clean.x.x$pol_rating_0<-0
pred.df.clean.x.x$trust_rating_0<-0
pred.df.clean.x.x$trust_rating_0<-0
pred.df.clean.x.x$trad_rating_0<-0
pred.df.clean.x.x$ind_rating_0<-0
pred.df.clean.x.x$imp_rating_0<-0
pred.df.clean.x.x$ex_rating_0<-0

#no climate here, to have 10 facets later
rating.vars.0 <- c("birth_rating_0","birth_rating_1", "birth_rating_2", "birth_rating_3",
                   "grat_rating_0","grat_rating_1", "grat_rating_2", "grat_rating_3",
                   "dep_rating_0","dep_rating_1", "dep_rating_2", "dep_rating_3",
                   "sat_rating_0","sat_rating_1", "sat_rating_2", "sat_rating_3",
                   "pol_rating_0","pol_rating_1", "pol_rating_2", "pol_rating_3",
                   "trust_rating_0","trust_rating_1", "trust_rating_2", "trust_rating_3",
                  "trad_rating_0", "trad_rating_1", "trad_rating_2", "trad_rating_3",
                   "ind_rating_0","ind_rating_1", "ind_rating_2", "ind_rating_3",
                   "imp_rating_0","imp_rating_1", "imp_rating_2", "imp_rating_3",
                   "ex_rating_0","ex_rating_1", "ex_rating_2", "ex_rating_3")

rating.vars <- c("cli_rating_1", "cli_rating_2", "cli_rating_3","birth_rating_1", "birth_rating_2", "birth_rating_3","grat_rating_1", "grat_rating_2", "grat_rating_3","dep_rating_1", "dep_rating_2", "dep_rating_3","sat_rating_1", "sat_rating_2", "sat_rating_3","pol_rating_1", "pol_rating_2", "pol_rating_3","trust_rating_1", "trust_rating_2", "trust_rating_3","trad_rating_1", "trad_rating_2", "trad_rating_3","ind_rating_1", "ind_rating_2", "ind_rating_3","imp_rating_1", "imp_rating_2", "imp_rating_3","ex_rating_1", "ex_rating_2", "ex_rating_3")
#Igor reordered so that it fits the graphic presentation below.


pred.df.clean.x.x.lng <- pred.df.clean.x.x %>%
                            gather(key = "Variable", value = "Scores", rating.vars) %>%
                            separate(Variable, into = c("Var", "Rating", "Time"), sep = "_")

pred.df.clean.x.x.lng.0 <- pred.df.clean.x.x %>%
                            gather(key = "Variable", value = "Scores", rating.vars.0) %>%
                            separate(Variable, into = c("Var", "Rating", "Time"), sep = "_")

```

## Graph

```{r graph.ggplot}

#analysis to get estimated means
test<-lmer(Scores~Time*Var+(1|ResponseId),data=pred.df.clean.x.x.lng)
library(emmeans)
summary(test)
emmeans(test, ~Var*Time)
#emmeans do not quite match the scores beloW!

descriptives<-as.data.frame(emmeans(test, ~Var*Time))

descriptives$Var <- factor(descriptives$Var , levels=c("ex", "imp", "ind", "trad", "trust", "pol","sat","dep","grat","birth","cli"))

descriptives$Var<-factor(descriptives$Var, labels =  c("Explicit \nPrejudice", "Implicit \nPrejudice", "Individualism", "Traditionalism", "Generalized \nTrust","Political \nPolarization","Life \nSatisfaction","Depression","Delay \nGratification","Birth Rate", "Concern about \nClimate Change"))


ggplot(descriptives, aes(x = Time, y = emmean, colour = Var, fill = Var, group = Var)) + 
  facet_wrap(. ~ Var, ncol = 3) + 
  geom_point(size = 1.5, show.legend = T, position = position_dodge(0)) + 
  geom_line(size = 0.5, show.legend = F, position = position_dodge(0)) +
  geom_errorbar(aes(ymin=asymp.LCL, ymax=asymp.UCL), size = 0.3, position = position_dodge(0), show.legend = F, width = 0.05) +   scale_y_continuous(breaks = seq(-20, 20, 10)) + 
  scale_x_discrete(labels = c("6 \nMonths", "1 \nYear", "2 \nYears")) + 
  coord_cartesian(ylim = c(-20, 20)) + 
  labs(x = "Timeline", y = "% Change", colour = "Issue", fill = "Issue") + 
  ggpubr::theme_pubclean(base_size = 12) + 
  theme_minimal(base_size = 12) +
  theme(axis.line = element_line(colour = "gray"), legend.position = "top") +
  theme(legend.position = "none")


#test with time as a continuous variable in month
pred.long<-pred.df.clean.x.x.lng.0
pred.long$Timeline[pred.long$Time=="0"]<-1
pred.long$Timeline[pred.long$Time=="1"]<-2
pred.long$Timeline[pred.long$Time=="2"]<-3
pred.long$Timeline[pred.long$Time=="3"]<-4

pred.long$Var <- factor(pred.long$Var , levels=c("ex", "imp", "ind", "trad", "trust", "pol","sat","dep","grat","birth"))

pred.long$Var<-factor(pred.long$Var, labels =  c("Explicit \nPrejudice", "Implicit \nPrejudice", "Individualism", "Traditionalism", "Generalized \nTrust","Political \nPolarization","Life \nSatisfaction","Depression","Delay of \nGratification","Birth Rate"))



library(ggthemes)

ggplot(pred.long, aes(x = Timeline, y = Scores, colour = Var)) + geom_boxplot(inherit.aes = F, aes(x = Time, y = Scores, colour = Var,alpha = 0.5))+
  #geom_point( alpha = 0.05) +
  geom_smooth(method = "loess")+  facet_wrap(. ~ Var, ncol = 5) + theme_minimal() +
  theme(axis.line = element_line(colour = "gray"), legend.position = "top") +
  theme(legend.position = "none")+scale_x_discrete(name="Timeline", labels = c("0"="Now","1"="6 Months","2"="Year","3"="2 Years")) +
  scale_y_continuous(name="% Change", limits=c(-50, 50))





```

